---
title: "lab10"
author: "Derek Maldonado"
date: "5/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ch 20
From Chapter 20, reproduce the bat and hat example (20.2) in R. Your code should represent the data in long-form, conduct the ANOVA, and report the ANOVA table. You will know if you did it correctly if you can reproduce the ANOVA table from the textbook.


```{r}
library(tibble)
library(ggplot2)

data20 <- tribble(~subject, ~Phonological_Similarity, ~age, ~num_correct,
        "s1","b1","a1",15,
        "s2","b1","a1",23,
        "s3","b1","a1",12,
        "s4","b1","a1",16,
        "s5","b1","a1",14,
        "s1","b2","a1",13,
        "s2","b2","a1",19,
        "s3","b2","a1",10,
        "s4","b2","a1",16,
        "s5","b2","a1",12,
        "s6","b1","a2",39,
        "s7","b1","a2",31,
        "s8","b1","a2",40,
        "s9","b1","a2",32,
        "s10","b1","a2",38,
        "s6","b2","a2",29,
        "s7","b2","a2",15,
        "s8","b2","a2",30,
        "s9","b2","a2",26,
        "s10","b2","a2",30,
        )

ggplot(data20,
       aes(y=num_correct,
           x=age,
           group=Phonological_Similarity,
           linetype=Phonological_Similarity))+
  geom_line()+
  geom_point()

aov20 <- aov(num_correct ~ Phonological_Similarity*age + Error(subject/Phonological_Similarity), data20)

summary(aov20)



#90% took the data from the class github because I didn't feel like typing it all out, also I don't know how to fix my graph to just show the 2 lines and not the vertical lines
```

## Ch 21
From Chapter 21, reproduce the phonological similarity example (21.2.1) in R. Your code should represent the data in long-form, conduct the ANOVA, and report the ANOVA table. You will know if you did it correctly if you can reproduce the ANOVA table from the textbook.


```{r}

subject <- rep(c("s1","s2","s3","s4"), each=10)
typB <- rep(rep(c("b1","b2"), each=5),4)
faceA <- rep(c("a1","a2","a3","a4","a5"),8)
centiseconds <- c(20,22,25,24,19,37,37,43,48,45,
                   9,8,21,21,21,34,35,35,37,39,
                   18,20,18,21,33,35,39,39,37,40,
                   5,14,16,22,23,38,49,51,50,52)

table21 <- tibble(subject,typB,faceA,centiseconds)

aov21 <- aov(centiseconds  ~ (subject + 
                                faceA%in%typB + 
                                typB + 
                                typB:subject),
               data = table21)


summary(aov21)


#25% struggled remembering the path from class for this one, did not seek out quasi_F
```



