---
title: "lab1"
author: "Derek Maldonado"
date: "2/4/2021"
bibliography: BCStats.bib
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QALMRI

for the paper by @slamecka_retroactive_1960


Q
  - Does learning new things affect the already learned information within one's mind
  -How does learning through original learning (OL) and interpolated learning (IL) practices affect the retroactive inhibition of read prose literature.
  
A
  - Remembering a learned prose passage is directly correlated by the amount of OL and inversely correlated by the amount of IL
  - There is no difference in learning between OL and IL
  
L
  - if the hypothesis is correct, there will be more retention with more OL and less retention with more IL
  - If the hypothesis is incorrect, there will be no difference in retention based on the amount of OL and IL done during the experiment
  
M
  - Variables
    IV - The number of OL and IL trials done by each subject
    DV - The subject's total correct anticipations
  - Subjects
    36 students in a general psych course in 4 40 minute sessions
  - material
    8 20 word sentences, all taken from the same book source
    sentences grouped into 4 pairs, 3 of them being used in the experiment
  - procedure
    each word exposed to subject for 3 seconds
    first pair used to practice OL and IL learning
    following pairs used in different ways as part of experiment
    all learning took total of 60 seconds
    
R
  - Descriptive
    Statistically significant F values found by way of ANOVA test
      Within only OL acquisition compared to OL x IL trials (F = 608.67)
      OL recall and # of IL trials (F = 28.06)
      OL recall only compared to OL x IL trials (F = 41.85)
  - Inferential
    The amount of words learned/retained through OL acquisition decreased with more IL trials, showing an inverse pattern. The amount of words learned/retained through IL acquisition increased with more IL trials 
    
I
  - Inferences
    with OL learning, there was an increase in retention based on the number of OL trials competed, and an inverse amount of retention based on the number of IL trails done when OL acquisition was used.
    There was an increase in retention based on the number of IL trials when IL acquisition was used.
    This allows the experimenter to reject the null hypothesis partially.
  - Flaws and Limitations
  The subjects used all came from the same class, therefore this is a small and specific grouping, not one that can be used to classify everyone.
    Psychology students may have a higher affinity to learning prose sentences compared to those in a more number-based STEM learning group.

## DATAFRAME
Create a dataframe that capable of representing the design, in as complete a way as you can, in the Slamecka (1960) paper.


```{r}
library(dplyr)

n <- 36
OL <- 3
IL <- 3
OLtrials <- c(2,4,8)
IL1 <- c(0,4,8)
IL2 <- c(4,8)

pt2 <- matrix(4, ncol=3, nrow=3)
colnames(pt2) <- c(OLtrials)
row.names(pt2) <- c(IL1)

subjects <- 1:36
OLsub <- rep(c(OLtrials),12)
ILsub <- rep(c(IL1),12)
recall <- sample(0:20,36, replace = TRUE)

longwise <- tibble(subjects,
                   OLsub,
                   ILsub,
                   recall)

mixed_design <- tribble(
  ~Subjects, ~OL, ~IL,
  1, 2, 0,
  1, 4, 4,
  1, 8, 8,
  2, 4, 0,
  2, 8, 4,
  2, 2, 8,
  3, 8, 0,
  3, 2, 4,
  3, 4, 8,
  4, 2, 4,
  4, 4, 0,
  4, 8, 8,
  5, 4, 4,
  5, 2, 8,
  5, 8, 0,
  6, 8, 4,
  6, 4, 8,
  6, 2, 0,
  7, 2, 8,
  7, 4, 0,
  7, 8, 4,
  8, 4, 8,
  8, 2, 4,
  8, 8, 0,
  9, 8, 8,
  9, 4, 4,
  9, 2, 0
)


full_mixed_design <- rbind(mixed_design,
                           mixed_design,
                           mixed_design,
                           mixed_design) %>%
  mutate(Subjects = rep(1:36, each=3),
         recall = sample(0:20,36*3, replace=TRUE))

full_mixed_design %>%
  group_by(OL,IL) %>%
  summarize(mean_recall = mean(recall), .groups="drop")


OL2_mean <- c(18.00,17.42,16.33)
OL4_mean <- c(40.33,40.17,44.50)
OL8_mean <- c(105.17,107.17,97.33)
OL2sd <- c(5.00, 6.38, 5.57)
OL4sd <- c(12.88, 8.69, 11.30)
OL8sd <- c(21.16, 22.12, 22.24)


widework <- data.frame(IL_trails = IL1,
                       OL2mean = OL2_mean,
                       SD = OL2sd,
                       OL4mean = OL4_mean,
                       SD = OL4sd,
                       OL8mean = OL8_mean,
                       SD = OL8sd)


#25% it was a struggle, as advertised. I especially fell apart near the end

```

## References
