---
title: "lab7"
author: "Derek Maldonado"
date: "4/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
Explain the concept of main effects and interactions with an example using R. For example, this could include a definition of main effects and interactions and a figure depicting main effects and an interaction along with an explanation of the patterns for each. A major point of this problem is for you to to engage in the task of developing an explanation of these concepts that would 1) be helpful for you to understand the concepts, and 2) could be helpful for others to understand these concepts.

```{r}
library(tibble)
library(ggplot2)
library(dplyr)
library(patchwork)

n <- 20

#the v vectors make up the raw data and function as the output measurement that has been affected by our iv1 and iv2 
v1 = c(11,9,7,11,12,7,12,11,10,10)
v2 = c(12,12,7,9,9,10,12,10,7,12)
v3 = c(17,20,22,13,21,16,23,19,20,19)
v4 = c(32,31,27,30,29,30,33,25,25,28)

#this creates a three column table showing which values from the v vectors are part of the a sample and which are part of the b sample
q1data <- tibble(iv1 = factor(rep(c("a1","a2"), each = n)),
                  iv2 = factor(rep(c("b1","b2"), n)),
                  dv = c(v1,v2,v3,v4))

#here we run a two-factor anova using the data from above then examine the value tables
a1 <- aov(dv ~ iv1*iv2, data = q1data)
model.tables(a1, type = "means")

#we create 3 graphs to show the means for the input values
A <- q1data %>%
  group_by(iv1) %>%
  summarise(mean_DV = mean(dv)) %>%
  ggplot(aes(y=mean_DV, x=iv1))+
   geom_bar(stat="identity", position="dodge") +
   ggtitle("Main effect A")

B <- q1data %>%
  group_by(iv2) %>%
  summarise(mean_DV = mean(dv)) %>%
  ggplot(aes(y=mean_DV, x=iv2))+
   geom_bar(stat="identity", position="dodge") +
   ggtitle("Main effect B")

AB <- q1data %>%
  group_by(iv1,iv2) %>%
  summarise(mean_DV = mean(dv)) %>%
  ggplot(aes(y=mean_DV, x=iv1, fill=iv2))+
   geom_bar(stat="identity", position="dodge")+
   ggtitle("AxB Interaction")

C <- ggplot(q1data, aes(x=iv1, y=dv, group=iv2, linetype= iv2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+

(A+B)/AB

#we create the 2x2 matrix for the means so that any effects can be clearly visible
tabling <- matrix(c(mean(v1),mean(v2),mean(v3),mean(v4)), ncol=2)
colnames(tabling) <- c('a1','a2')
rownames(tabling) <- c('b1','b2')
table1 <- as.table(tabling)

#print out the table as well as a line graph showing a simple version of the effects from the IVs
C
table1

#75%
```
as we can see because of these graphs, there is a significant main effect from iv1 and a minimal one coming from iv2, also there is a strong interaction effect between iv1 and iv2.

## Q3
n chapter 10 of Crump et al. (2018), there is a discussion of patterns of main effects and interactions that can occur in a 2x2 design, which represents perhaps the simplest factorial design. There are 8 possible outcomes discussed https://crumplab.github.io/statistics/more-on-factorial-designs.html#looking-at-main-effects-and-interactions. Examples of these 8 outcomes are shown in two figures, one with bar graphs, and one with line graphs. Reproduce either of these figures using ggplot2

```{r}
library(ggplot2)
library(patchwork)

#graph 1 no effects
a <- c(5,5,5,5,5)
a2 <- c(5,5,5,5,5)
b <- c(5,5,5,5,5)
b2 <- c(5,5,5,5,5)
g1data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
A <- ggplot(g1data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("No effects")+
  theme(plot.title = element_text(face = "bold"))

#graph 2 Interaction effect
a <- c(10,10,10,10,10)
a2 <- c(5,5,5,5,5)
b <- c(5,5,5,5,5)
b2 <- c(10,10,10,10,10)
g2data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
B <- ggplot(g2data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("Interaction effect")+
  theme(plot.title = element_text(face = "bold"))

#graph 3 IV2 main effect
a <- c(10,10,10,10,10)
a2 <- c(10,10,10,10,10)
b <- c(5,5,5,5,5)
b2 <- c(5,5,5,5,5)
g3data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
C <- ggplot(g3data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("IV2 main effect")+
  theme(plot.title = element_text(face = "bold"))

#graph 4 IV1 main effect
a <- c(10,10,10,10,10)
a2 <- c(5,5,5,5,5)
b <- c(10,10,10,10,10)
b2 <- c(5,5,5,5,5)
g4data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
D <- ggplot(g4data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("IV1 main effect")+
  theme(plot.title = element_text(face = "bold"))

#graph 5 IV1 main + Interaction
a <- c(15,15,15,15,15)
a2 <- c(10,10,10,10,10)
b <- c(0,0,0,0,0)
b2 <- c(5,5,5,5,5)
g5data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
E <- ggplot(g5data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("IV1 main + Interaction")+
  theme(plot.title = element_text(face = "bold"))

#graph 6 IV2 main + interaction
a <- c(0,0,0,0,0)
a2 <- c(15,15,15,15,15)
b <- c(5,5,5,5,5)
b2 <- c(10,10,10,10,10)
g6data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
G <- ggplot(g6data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("IV2 main + Interaction")+
  theme(plot.title = element_text(face = "bold"))

#graph 7 IV1 main + IV2 main
a <- c(5,5,5,5,5)
a2 <- c(10,10,10,10,10)
b <- c(10,10,10,10,10)
b2 <- c(15,15,15,15,15)
g7data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
H <- ggplot(g7data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("IV1 main + IV2 main")+
  theme(plot.title = element_text(face = "bold"))

#graph 8 all effects
a <- c(10,10,10,10,10)
a2 <- c(20,20,20,20,20)
b <- c(5,5,5,5,5)
b2 <- c(7,7,7,7,7)
g8data <- tibble(means = c(a,a2,b,b2),
                  IV1 = rep(c("A","B"), each = 10),
                  IV2 = rep(rep(c("1","2"), each = 5), 2)
                  )
I <- ggplot(g8data, aes(x=IV1, y=means, group=IV2, linetype= IV2))+
     geom_point(stat = "summary", fun="mean")+
     geom_line(stat = "summary", fun="mean")+
  expand_limits(y=c(0,20))+
  ggtitle("IV1 main + IV2 main + interaction")+
  theme(plot.title = element_text(face = "bold"))

#graphs together
(A+B+C+D)/(E+G+H+I)

#100%

```

