---
title: "lab8"
author: "Derek Maldonado"
date: "5/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Consider a 2x2 design. Assume the DV is measured from a normal distribution with mean 0, and standard deviation 1. Assume that the main effect of A causes a total shift of .5 standard deviations of the mean between the levels. Assume that level 1 of B is a control, where you expect to measure the standard effect of A. Assume that level 2 of B is an experimental factor intended to reduce the effect of A by .25 standard deviations

create a ggplot2 figure that depicts the expected results from this design


```{r}
library(tibble)
library(ggplot2)

A <- c(0,0.5)
B <- c(0,0)
AB <- c(0,0,0,0.25)
gmean <- 0

ansdata <- tibble()
for(i in 1:length(A)){
  for(j in 1:length(B)){
    IVA <- i
    IVB <- j
    DV <- gmean+A[i]+B[j]+AB[(i-1)*length(B)+j]
    scoregm <- gmean
    scoreA <- A[i]
    scoreB <- B[j]
    scoreAB <- AB[(i-1)*length(B)+j]
    row_entry <- tibble(IVA,IVB,DV,scoregm,scoreA,scoreB,scoreAB)
    ansdata <- rbind(ansdata,row_entry)
  }
}

knitr::kable(ansdata)

liner <- ggplot(ansdata,
                aes(y=DV,
                    x=IVA,
                    linetype=as.factor(IVB)))+
  geom_line()+
  geom_point()

liner

```

How many subjects are needed to detect the main effect of A with power = .8?

```{r}
library(tibble)
library(ggplot2)

N <- 55

Ap <- c()
Bp <- c()
ABp <- c()

for(i in 1:1000){
  IVA <- rep(rep(c("1","2"), each=2),N)
  IVB <- rep(rep(c("1","2"), 2),N)
  DV <- c(replicate(N,c(rnorm(1,0,1),
                        rnorm(1,0,1),
                        rnorm(1,.5,1),
                        rnorm(1,.25,1) 
          )))
  sim_df <- data.frame(IVA,IVB,DV)

  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  Ap[i]<-aov_results[[1]]$`Pr(>F)`[1]
  Bp[i]<-aov_results[[1]]$`Pr(>F)`[2]
  ABp[i]<-aov_results[[1]]$`Pr(>F)`[3]
}

length(Ap[Ap<0.05])/1000
#length(Bp[Bp<0.05])/1000
#length(ABp[ABp<0.05])/1000

```
Need 56 particpants for 0.8 power on main effect of A



How many subjects are needed to detect the interaction effect with power = .8?
```{r}
N <- 500


ABp <- c()

for(i in 1:1000){
  IVA <- rep(rep(c("1","2"), each=2),N)
  IVB <- rep(rep(c("1","2"), 2),N)
  DV <- c(replicate(N,c(rnorm(1,0,1),
                        rnorm(1,0,1),
                        rnorm(1,.5,1),
                        rnorm(1,.25,1) 
          )))
  sim_df <- data.frame(IVA,IVB,DV)

  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  ABp[i]<-aov_results[[1]]$`Pr(>F)`[3]
}


length(ABp[ABp<0.05])/1000

```
Will need about 500 participants to obtain a power of 0.8